fGFPCA dynamic prediction

### 2022-11-11 f-GFPCA.Rmd
update simulation scheme

## Toy simulation with a very simple function

@@

- Below we generate 100 latent functions and their corresponding binary functions

$$Z_i(t) = a_{i1}sin^2(t)+a_{i2}cos^2(t)+a_{i3}t^2+a_{i4}t$$

$$Y_i(t) \sim Binomial(\frac{exp(Z_i(t))}{1+exp(Z_i(t))})$$

```{r}
## four subjects
ggplot(df_train2 %>% filter(id %in% 15:18))+
  geom_point(aes(x=argvals, y=Y, col = "blue"),  size = 0.1,
             show.legend = F)+
  geom_line(aes(x=argvals, y=Z, col = "Red"), 
            show.legend = F)+
  facet_wrap(~id)+
  labs(title = "Simulated function for the 4 subjects",
       x = "time", y ="")

## all subjects latent functions
ggplot(df_train2)+
  geom_line(aes(x=argvals, y=Z, col = as.factor(id)), 
            show.legend = F)+
  labs(title = "Simulated latent functions for all subjects")

```

- Below is the predicted latent functions from local GLMM models

```{r}

ggplot(df_pred_latent %>% filter(id %in% 15:18))+
  geom_line(aes(x = argvals, y=Zhat))+
  geom_line(aes(x=argvals, y = Z, col = "red"))+
  geom_point(aes(x=argvals, y = Y, col = "blue"), size = 0.01)+
  facet_wrap(~id)

```

>>>>

## Toy simulation

### Simulation set-up

@@

- For a random sample i:

$$\eta_i(t) =f_0(t)+ \xi_{i1}sin(2\pi t)+\xi_{i2}cos(2\pi t)+\xi_{i3}sin(4\pi t)+\xi_{i4}cos(4\pi t)$$

where $\boldsymbol{\xi}_{i} \sim N(0, \boldsymbol{\Gamma})$

$$Y_i(t) \sim Binomial(\frac{exp(\eta_i(t))}{1+exp(\eta_i(t))})$$

In this simulation, we set $f_0(t) = 0$. In the fPCA framework, $\Gamma$ is a diagonal matrix and diagonal elements correponds to eigen values.

```{r}
df %>% 
  filter(id %in% 1:4) %>% 
  ggplot()+
  geom_line(aes(x=sind, y=eta_i))+
  geom_point(aes(x=sind, y=Y), size = 0.05)+
  facet_wrap(~id)+
  labs(title = "Four simulated subjects")
```

### Local GLMMS

```{r}
source(here("Code/GLMM-FPCA.R"))
```

```{r}
par(mfrow=c(3, 2))
## mean functions
plot(mid, fpca_fit$mu, ylab = "mean")

## eigenfunctions
plot(mid, fpca_fit$efunctions[, 1], ylab='PC1')
plot(mid, fpca_fit$efunctions[, 2], ylab='PC2')
plot(mid, fpca_fit$efunctions[, 3], ylab='PC3')
plot(mid, fpca_fit$efunctions[, 4], ylab='PC4')
```

```{r}
ggplot(df_pred_latent %>% filter(id %in% 1:4))+
  geom_line(aes(x = sind_inx, y=eta_hat))+
  geom_line(aes(x=sind_inx, y = eta_i, col = "red"))+
  geom_point(aes(x=sind_inx, y = Y, col = "blue"), size = 0.01)+
  facet_wrap(~id)+
  labs(title = "Predicted latent functions for four subjects")
```

### Prediction

Assume we have a new observation with outcome $\boldsymbol{Y}$ up to $t_m$. If we have an Gaussian function, the scores of this subject can be estimated with empirical Bayes:

$$\hat{\boldsymbol{\xi}} = E(\boldsymbol{\xi}|\boldsymbol{Y}) =  \boldsymbol{\hat{\Gamma}\Phi ^T}(\boldsymbol{\Phi\hat{\Gamma}\Phi^T}+\hat{\sigma_{\epsilon}}^2\boldsymbol{I_m})^{-1}\boldsymbol{Y}$$
Where $\hat{\boldsymbol{\Gamma}}$ is the diagonal matrix of eigenvalues, $\boldsymbol{\Phi}$ is the matrix of eigenfunctions. These should be constant across all subjects.

However, for binary data with latent function, there are a couple of differences:

1. We do not have a $\sigma_{\epsilon}^2$ in the model formula

2. We do not have an Gaussian outcome $\boldsymbol{Y}$. The $\boldsymbol{Y}$ should be replaced with the values of latent function $\eta(t)$. However, we also could not observe the value of $\eta(t)$.

Two options I am thinking of:

1. If we'd like to make predictions on more than one sample, we may be able to fit GLMM on test samples and predict $\eta(t)$
2. Do a different set of derivation with MLE+EB like [this paper](https://www.jstor.org/stable/pdf/2533715.pdf?refreqid=excelsior%3A135dd765c41e4b46eb39bb6ef7a71053&ab_segments=&origin=&acceptTC=1)

Another thing I have been thinking of is how to extend the grid of eigenfunctions from binned-grid to full-grid. I am using interpolation now, but only because I am not sure how to do it with basis functions. It looks like we will need K more sparse PCA, but it will introduce a lot more scores.
